Classless Java: Tuning Java Interfaces

### 2015-12-06 paper/sections/Formalization.tex
Section 2 revied
Section 3 reviewd

\section{Interaction of Interface Methods with Interface Composition}
Before formalizing Classless Java, it is helpful to informally discuss the behaviour
of methods in Java 8 interfaces.

@@

\subsection{Classifying Outcomes of Interface Composition}
When interfaces are composed and methods with the same name (and
signature) exist, there are 3 possible outcomes.
We illustrate the possible outcomes using the following (correct) declarations:
\begin{lstlisting}
interface A1{T m(); }
interface A2 extends A1{default T m(){ ... } }
interface A3 extends A2{T m(); }

@@

\noindent
What happens if a new interface \Q@M@ extends one \Q@A@${}_i$ and one
\Q@B@${}_j$? Nine representative cases are shown next:\\*
\noindent
\begin{tabular}{|l|l|l|l|}
\hline
\textbf{M extends} & \textbf{A1}                  & \textbf{A2} & \textbf{A3} \\ \hline
\textbf{B1}        & conservative error                        & conflict error      & conservative error       \\ \hline
\textbf{B2}        & both abstract, accepted                        & conservative error       & both abstract, accepted       \\ \hline
\textbf{B3}        & \textbf{conservative error} &conflict  error       & conservative error      \\ \hline
\end{tabular}
\\*
We try to classify the results in the table:
\begin{itemize}
\item \textbf{conflict error} happens when the methods from both are implemented, which is also an error in most trait models.
\item \textbf{both abstract, accepted} happens when the methods from both are abstract, which is also considered correct in all
  trait models.
\item \textbf{conservative error} happens when only one method is implemented
  (leaving another one abstract), which is different from what we would expect in
  a trait model, but is coherent with the conservative idea that a method
  defined in an interface should not silently satisfy a method in another one.
\end{itemize}

>>>>

\section{Interaction of Interface Methods with Interface Composition}
%Before formalizing Classless Java, it is helpful to informally discuss the behaviour
%of methods in Java 8 interfaces.

@@

\subsection{Classifying Outcomes of Interface Composition}
%When interfaces are composed and methods with the same name (and
%signature) exist, there are 3 possible outcomes.
%
We now try to classify possible outcomes for composition of methods with the same name (and
signature).
We will use the following (correct) declarations:
\begin{lstlisting}
interface A1{T m(); }
interface A2 extends A1{default T m(){ ... } }
interface A3 extends A2{T m(); }

@@

\noindent
What happens if a new interface \Q@M@ extends one \Q@A@${}_i$ and one
\Q@B@${}_j$?\\*
% Nine representative cases are shown next:\\*
\noindent
\begin{tabular}{|l|l|l|l|}
\hline
\textbf{M extends} & \textbf{A1}                  & \textbf{A2} & \textbf{A3} \\ \hline
\textbf{B1}        & conservative error                        & conflict error      & conservative error       \\ \hline
\textbf{B2}        & both abstract, accepted                        & conservative error       & both abstract, accepted       \\ \hline
\textbf{B3}        & \textbf{conservative error} &conflict  error       & conservative error      \\ \hline
\end{tabular}
\\*
%We try to classify the results in the table:
\begin{itemize}
\item \textbf{conflict error} happens when the methods from both are implemented, which is also an error in most trait models.
\item \textbf{both abstract, accepted} happens when the methods from both are abstract, which is also considered correct in all
  trait models.
\item \textbf{conservative error} happens when only one method is implemented
  (leaving another one abstract), which is different from what we would expect in
  a trait model, but is coherent with the conservative idea that a method
  defined in an interface should not silently satisfy a method in another one.
\end{itemize}
